---
title: "Course Project  - Practical Machine Learning"
author: "Andreia Vieira"
date: "12/08/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prediction Assignment Writeup

### 1) Summay

Company are realized measurements regularly about themselves  to improve their health, to find patterns in their behavior, or because they are tech geeks. In this perspective,participants were asked to perform barbell lifts correctly and incorrectly in 5 different ways. Based on this experience, the activity aims to predict the way in which new participants performed the exercises described and answer questions in the associated course questionnaire. The initial studies were based on the experience of six participants. The machine learning algorithm, which uses the class variable in the training set, is applied to the 20 test cases. The activity was developed in RStudio and published in html format.

### 2) Data

The training data for this project are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv


The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. 

VELLOSO, E.; BULLING, A.; GELLERSEN, H.; UGULINO, W.; FUKS, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013.

### 3) Data Loading 

```{r message=FALSE, warning=FALSE, , echo=TRUE}

library(caret)
#library(knitr)
library(ggplot2)
library(rpart)
#library(rpart.plot)
library(lares)
library(dplyr)
library(Hmisc)
library(corrplot)
#library(rattle)
library(randomForest)
library(gbm)
set.seed(2020)

# set the URL for the download
Train <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
Validation  <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

# download the datasets
training <- read.csv(url(Train))
validation  <- read.csv(url(Validation))

str(training)



```

### 4) Data cleaning

```{r , echo = TRUE, warning=FALSE, message=TRUE}

data_training <-training


# Variable characteristics
#str(data_training)


# remove variables that are mostly NA
data_NA    <- sapply(data_training, function(x) mean(is.na(x))) > 0.95
data_training <- data_training[, data_NA==FALSE]
dim(data_training)

# values close to zero - weak correlation

data_zero <- nearZeroVar(data_training)
data_training<- data_training[, -data_zero]
dim(data_training)

#remover variables init
data_training <- data_training[, -(1:5)]
dim(data_training)


# split

data_split <- createDataPartition(data_training$classe, p=0.7, list=FALSE)

TrainSet <- data_training[data_split,]
TestSet  <- data_training[-data_split,]

dim(TrainSet)
dim(TestSet)

#Filtered variables

str(TrainSet)


```
### 5) - Correlation Analysis
```{r , echo = TRUE, warning=FALSE, message=FALSE}


#  Dataset without target variable - 53 variables
Avaiable_Correlation<-TrainSet[, -54]

# Correlations all variables
Matrix_correlation<-cor(Avaiable_Correlation)
corrplot(Matrix_correlation, type="upper", order="hclust", tl.col="black", tl.srt=45, tl.cex = 0.5)



#Variables that showed some high correlations - 29 variables
corr_cross(Avaiable_Correlation)
corr_cross(Avaiable_Correlation, type = 2)
variables<-corr_cross(Avaiable_Correlation, plot=FALSE)



#filter
Newdataset<-select(Avaiable_Correlation,accel_arm_z, yaw_belt, accel_forearm_y, magnet_dumbbell_x,
total_accel_dumbbell,yaw_dumbbell,magnet_arm_y,roll_belt,accel_arm_x,pitch_dumbbell,
roll_belt,magnet_arm_x,gyros_belt_x,accel_belt_z,magnet_belt_y,magnet_arm_z,accel_belt_z,
magnet_forearm_y,magnet_dumbbell_y,accel_dumbbell_y,roll_belt,roll_belt,total_accel_belt,
pitch_belt,accel_belt_y,total_accel_belt,roll_belt,gyros_arm_x,accel_belt_x,pitch_belt,
accel_dumbbell_z,magnet_arm_z,yaw_belt,magnet_arm_x,accel_dumbbell_x,accel_arm_y,magnet_arm_y,
magnet_dumbbell_y,accel_arm_y,magnet_belt_z,accel_belt_z,total_accel_belt,accel_belt_z,
accel_belt_x,accel_belt_z,accel_belt_y,accel_belt_y,gyros_arm_y,magnet_belt_x,magnet_belt_x)


# Correlations selected variables - 29 variables
Matrix_correlation<-cor(Newdataset)
corrplot(Matrix_correlation, type="upper", order="hclust", tl.col="black", tl.srt=45, tl.cex = 0.5)

   

```
### 6) Prediction Model Building

#### Build two different models: glm and rf

#### 6.1 Method: gbm and rf- training

```{r , echo = TRUE, warning=FALSE, message=FALSE}


#gbm model
set.seed(2020)
model_gbm <-train(classe ~ ., method="gbm", data=TrainSet, trControl= trainControl(method= "repeatedcv", number = 5, repeats = 1), verbose = FALSE)
model_gbm$finalModel

#rf model
set.seed(2021)
model_rf <- train(classe ~ ., data=TrainSet, method="rf",
                          trControl=trainControl(method="cv", number=5, verboseIter=FALSE))
model_rf$finalModel



```
#### 6.2 Method: gbm and rf- test

```{r , echo = TRUE, warning=FALSE, message=FALSE}

# test model gbm
predict_gbm <- predict(model_gbm, newdata=TestSet)
confMat_gbm  <- confusionMatrix(table(predict_gbm, TestSet$classe)) 
confMat_gbm

# test model rf
predict_rf <- predict(model_rf, newdata=TestSet)
confMat_rf  <- confusionMatrix(table(predict_rf, TestSet$classe)) 
confMat_rf




```
#### 6.3 Method: gbm and rf- plot
```{r , echo = TRUE, warning=FALSE, message=FALSE}

# plot matrix gbm results
plot(confMat_gbm$table, col = confMat_gbm$byClass, 
    main = paste("gbm - Accuracy =", round(confMat_gbm$overall['Accuracy'], 4)))

# plot matrix rf results
plot(confMat_rf$table, col = confMat_rf$byClass, 
    main = paste("rf - Accuracy =", round(confMat_rf$overall['Accuracy'], 4)))



```

### 7) Applying the Selected Model (best model) to the Test Data (Validation)

```{r , echo = TRUE, warning=FALSE, message=FALSE}

#best result  method rf
predict_validation <- predict(model_rf, newdata=validation)
predict_validation

```
  
  


```

  
  